---
title: "sdc: Swiss Data Cube"
output: 
  github_document:
     html_preview: true
editor_options: 
  chunk_output_type: console
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(collapse = T, tidy=T, comment = NA, warning=F, message=F)
```

## Installation

You can install sdc from github with:

```{r, eval=F}
# Install remotes if not available
if(!"remotes" %in% installed.packages()[,"Package"]) install.packages("remotes")

# Install sdc package from Github
remotes::install_github("RS-eco/sdc", build_vignettes = T)
```

If package installation with `install_github()` fails, try this as an alternative:

```{r, eval=F}
tmp_zip <- tempfile(fileext = ".zip")
source_url <- "https://api.github.com/repos/RS-eco/bavDC/zipball/main"
utils::download.file(source_url, destfile=tmp_zip, method="wget")
file.exists(tmp_zip)

remotes::install_local(tmp_zip)
```

After installation, simply load the sdc package:

```{r}
library(sdc)
```

**If you encounter a bug or if you have any problems, please file an [issue](https://github.com/RS-eco/sdc/issues) on Github.**

## Datasets

### GADM data

Shapefiles of Canton Glarus and Switzerland can be accessed by:

```{r}
data("glarus")
data("che")
```

### CH2018 data

The CH2018 climate data for Switzerland can be accessed by:

```{r}
data("ch2018_bioclim_che")
```

### Euro-Cordex Data

The bioclimatic Euro-Cordex data can be accessed by:

```{r}
# Euro-Cordex data for Switzerland
data("cordex_bioclim_che")
```

**Note:** Please also have a look at the corresponding [vignette](https://github.com/RS-eco/sdc/blob/main/vignettes/sdc-euro-cordex.Rmd) for detailed information about the dataset and the related variables.

### Corine Land-cover data

```{r}
# Corine data for Switzerland
data("corine_lc_che")
```

**Note:** Please also have a look at the corresponding [vignette](https://github.com/RS-eco/sdc/blob/main/vignettes/sdc-landcover.Rmd) for detailed information about the dataset and the related variables.

### SRTM data

The SRTM elevation data can be accessed by:

```{r}
# SRTM data for Glarus
data("srtm3_glarus_1arc")

# SRTM data for Switzerland
data("srtm_csi_che_3arc")
data("srtm3_che_3arc")

# High resolution SRTM data for whole of Switzerland
load(system.file("extdata", "srtm3_che_1arc.rda", package = "sdc"))
```

**Note:** Please also have a look at the corresponding [vignette](https://github.com/RS-eco/sdc/blob/main/vignettes/sdc-elevation.Rmd) for detailed information about the dataset and the related variables.
